---
title: "Follow-up question"
author: "Jinlong"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```         
```

> 

```{r}
options(device = "png")
df_interaction <- read.csv("police_killings.csv", fileEncoding = "latin1") %>%
  clean_names() %>%
  mutate(
    comp_income = as.numeric(comp_income),
    pov = as.numeric(pov),
    urate = as.numeric(urate),
    college = as.numeric(college)
  )

df_interaction <- df_interaction %>%
  mutate(
    income_group = case_when(
      comp_income < 0.8 ~ "Low income",
      comp_income <= 1.2 ~ "Medium income",
      comp_income > 1.2 ~ "High income",
      TRUE ~ NA_character_
    ),
    income_group = factor(
      income_group,
      levels = c("Low income", "Medium income", "High income")
    )
  )

overall_factor_analysis <- function(data, var, var_label) {
  
  df_sub <- data %>%
    filter(!is.na(income_group), !is.na(.data[[var]]))
  
  cut1 <- quantile(df_sub[[var]], 1/3, na.rm = TRUE)
  cut2 <- quantile(df_sub[[var]], 2/3, na.rm = TRUE)
  
  df_sub <- df_sub %>%
    mutate(
      factor_group = case_when(
        .data[[var]] < cut1 ~ paste("Low", var_label),
        .data[[var]] < cut2 ~ paste("Medium", var_label),
        TRUE ~ paste("High", var_label)
      ),
      factor_group = factor(
        factor_group,
        levels = c(
          paste("Low", var_label),
          paste("Medium", var_label),
          paste("High", var_label)
        )
      )
    )
  
  overall_table <- df_sub %>%
    count(income_group, factor_group) %>%
    mutate(
      pct_total = round(n / sum(n) * 100, 1)
    ) %>%
    arrange(desc(n))
  
  count_table <- table(df_sub$income_group, df_sub$factor_group)
  chi_result <- chisq.test(count_table)
  
  p_label <- ifelse(
    chi_result$p.value < 0.001,
    "p < 0.001",
    paste0("p = ", round(chi_result$p.value, 4))
  )
  
  p <- ggplot(overall_table, aes(x = factor_group, y = income_group, fill = pct_total)) +
    geom_tile(color = "white", linewidth = 0.8) +
    geom_text(aes(label = paste0(pct_total, "%")), size = 5) +
    scale_fill_gradient(low = "lightblue", high = "darkblue") +
    labs(
      title = paste("Overall distribution: Income x", var_label, "(", p_label, ")"),
      subtitle = "Share of total incidents",
      x = paste(var_label, "group"),
      y = "Income group",
      fill = "% of total"
    ) +
    theme_minimal(base_size = 14)
  
  cat("\n===== Overall Table:", var_label, "=====\n")
  print(overall_table)
  
  cat("\n===== Chi-square Test:", var_label, "=====\n")
  print(chi_result)
  
  return(list(
    overall_table = overall_table,
    chi_result = chi_result,
    plot = p
  ))
}

# 4. Run for multiple factors

res_pov <- overall_factor_analysis(df_interaction, "pov", "poverty")
print(res_pov$plot)

res_urate <- overall_factor_analysis(df_interaction, "urate", "unemployment")
print(res_urate$plot)

res_college <- overall_factor_analysis(df_interaction, "college", "education")
print(res_college$plot)
```

The results provide strong evidence that certain socioeconomic factors amplify the effect of low income on police killings. Chi-square tests show highly significant associations between income and poverty, unemployment, and education (all p-values \< 0.001), indicating that these factors are not independent. More importantly, the overall distribution reveals that low income alone is insufficient to explain the concentration of incidents. For example, low-income tracts with high poverty account for 26.9% of all police killings, compared to only 3.4% for low-income tracts with low poverty. Similar amplification patterns are observed for unemployment (20.6% vs. 5.2%) and education (21.9% in low-income, low-education tracts vs. only 0.6% in high-income, low-education tracts). These large disparities demonstrate that police killings are most concentrated where multiple disadvantages co-occur. Therefore, the findings suggest a clear compounding effect: low income becomes significantly more impactful when combined with high poverty, high unemployment, or low education, rather than operating as a standalone factor
