---
title: "Police Killings - Armed Status Prediction"
author: "Group 14"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Load Libraries

```{r libraries}
library(tidyverse)
library(janitor)
library(ggplot2)
library(patchwork)     # Combine plots
library(pROC)          # AUC / ROC curve
library(caret)         # Confusion matrix
library(rpart)         # Decision tree
library(rpart.plot)    # Decision tree visualization
library(themis)        # SMOTE for class imbalance
library(recipes)       # Data preprocessing recipes
```

---

## 1. Data Loading

```{r load-data}
df <- read.csv("police_killings.csv")
```

---

## 2. Data Cleaning

```{r clean-names-and-whitespace}
df <- df |>
  clean_names() |>
  mutate(across(where(is.character), ~ str_squish(.))) |>
  mutate(across(where(is.character), ~ na_if(.x, "")))
```

```{r standardize-na}
df <- df |>
  mutate(across(where(is.character),
               ~ case_when(
                 .x %in% c("NA", "N/A", "null", "NULL", "?",
                            "unknown", "Unknown", "-", " - ", " -", "- ") ~ NA_character_,
                 TRUE ~ .x
               )))
```

```{r fix-types}
df <- df |>
  mutate(
    age            = as.integer(age),
    gender         = as.factor(gender),
    raceethnicity  = as.factor(raceethnicity),
    month          = as.factor(month),
    state          = as.factor(state),
    share_white    = as.numeric(share_white),
    share_black    = as.numeric(share_black),
    comp_income    = as.numeric(comp_income),
    share_hispanic = as.numeric(share_hispanic),
    p_income       = as.numeric(p_income),
    pov            = as.numeric(pov),
    urate          = as.numeric(urate),
    college        = as.numeric(college)
  )
```

```{r recode-armed-cause}
df <- df |>
  mutate(
    armed = case_when(armed == "No" ~ "No", TRUE ~ "Yes"),
    cause = case_when(cause == "Gunshot" ~ "Gunshot", TRUE ~ "Other")
  ) |>
  mutate(
    cause = as.factor(cause),
    armed = factor(armed, levels = c("No", "Yes"))
  )
```

---

## 3. Prepare Modeling Data

```{r prep-model-data}
armedpredict <- df |>
  select(armed, college, urate, h_income, pov) |>
  na.omit() |>
  mutate(
    armed    = factor(armed, levels = c("No", "Yes")),
    armed_yn = ifelse(armed == "No", 0, 1)
  )

feat_cols <- c("college", "urate", "h_income", "pov")

cat("Rows used for modeling:", nrow(armedpredict), "\n")
cat("\nClass distribution (original):\n")
print(table(armedpredict$armed))
cat("\nClass ratio (Yes:No):",
    round(sum(armedpredict$armed == "Yes") / sum(armedpredict$armed == "No"), 2), "\n")
```

---

## 4. Handling Class Imbalance with SMOTE

The dataset is imbalanced (~78% armed). SMOTE (Synthetic Minority
Oversampling Technique) generates synthetic samples for the minority
class ("No") to create a more balanced training set.

```{r smote}
set.seed(42)

smote_recipe <- recipe(armed ~ college + urate + h_income + pov,
                       data = armedpredict) |>
  step_smote(armed, over_ratio = 1)

smote_prep  <- prep(smote_recipe, training = armedpredict)
df_balanced <- bake(smote_prep, new_data = NULL) |>
  mutate(
    armed    = factor(armed, levels = c("No", "Yes")),
    armed_yn = ifelse(armed == "No", 0, 1)
  )

cat("Class distribution after SMOTE:\n")
print(table(df_balanced$armed))
```

```{r visualize-imbalance, fig.width=7, fig.height=4}
bind_rows(
  armedpredict |> count(armed) |> mutate(Dataset = "Original"),
  df_balanced  |> count(armed) |> mutate(Dataset = "After SMOTE")
) |>
  ggplot(aes(x = armed, y = n, fill = armed)) +
  geom_col(show.legend = FALSE) +
  geom_text(aes(label = n), vjust = -0.4) +
  facet_wrap(~ Dataset) +
  scale_fill_manual(values = c("No" = "#d73027", "Yes" = "#1a9850")) +
  labs(
    title = "Class Distribution Before and After SMOTE",
    x = "Armed", y = "Count"
  ) +
  theme_minimal()
```

---

## 5. Logistic Regression (trained on SMOTE data)

### 5.1 Fit Model

```{r fit-logistic}
armedmodel <- glm(
  armed_yn ~ college + urate + h_income + pov,
  data   = df_balanced,
  family = binomial
)

summary(armedmodel)
```

### 5.2 Evaluate on Original Data

```{r logistic-eval}
armedpredict$predicted_prob  <- predict(
  armedmodel,
  newdata = armedpredict[, feat_cols],
  type    = "response"
)
armedpredict$predicted_class <- factor(
  ifelse(armedpredict$predicted_prob >= 0.5, "Yes", "No"),
  levels = c("No", "Yes")
)

conf_mat <- confusionMatrix(
  data      = armedpredict$predicted_class,
  reference = armedpredict$armed,
  positive  = "Yes"
)
print(conf_mat)

roc_obj <- roc(armedpredict$armed_yn, armedpredict$predicted_prob)
auc_val <- auc(roc_obj)
cat("\nLogistic Regression AUC:", round(auc_val, 4), "\n")
```

### 5.3 ROC Curve

```{r logistic-roc}
ggroc(roc_obj, color = "steelblue", size = 1) +
  geom_abline(slope = 1, intercept = 1, linetype = "dashed", color = "grey50") +
  annotate("text", x = 0.3, y = 0.2,
           label = paste0("AUC = ", round(auc_val, 3)),
           size = 5, color = "steelblue") +
  labs(
    title = "ROC Curve - Logistic Regression (SMOTE)",
    x = "Specificity", y = "Sensitivity"
  ) +
  theme_minimal()
```

### 5.4 Predicted Probability Plots

```{r logistic-prob-plots, fig.width=10, fig.height=8}
p1 <- ggplot(armedpredict, aes(x = pov, y = predicted_prob)) +
  geom_point(alpha = 0.4) + geom_smooth() +
  labs(title = "Poverty Rate", x = "Poverty Rate", y = "Predicted Probability") +
  theme_minimal()

p2 <- ggplot(armedpredict, aes(x = college, y = predicted_prob)) +
  geom_point(alpha = 0.4) + geom_smooth() +
  labs(title = "College Attainment", x = "College Attainment", y = "Predicted Probability") +
  theme_minimal()

p3 <- ggplot(armedpredict, aes(x = urate, y = predicted_prob)) +
  geom_point(alpha = 0.4) + geom_smooth() +
  labs(title = "Unemployment Rate", x = "Unemployment Rate", y = "Predicted Probability") +
  theme_minimal()

p4 <- ggplot(armedpredict, aes(x = h_income, y = predicted_prob)) +
  geom_point(alpha = 0.4) + geom_smooth() +
  labs(title = "Household Income", x = "Household Income", y = "Predicted Probability") +
  theme_minimal()

(p1 + p2) / (p3 + p4) +
  plot_annotation(title = "Predicted Probability Victim Was Armed (Logistic Regression)")
```

---

## 6. Decision Tree (trained on SMOTE data)

### 6.1 Fit Decision Tree

```{r fit-tree}
set.seed(42)

tree_model <- rpart(
  armed ~ college + urate + h_income + pov,
  data    = df_balanced,
  method  = "class",
  control = rpart.control(minsplit = 10, cp = 0.01)
)

printcp(tree_model)
```

### 6.2 Visualize the Tree

```{r plot-tree, fig.width=10, fig.height=6}
rpart.plot(
  tree_model,
  type          = 3,
  extra         = 104,
  box.palette   = c("#d73027", "#1a9850"),
  fallen.leaves = TRUE,
  main          = "Decision Tree: Predicting Armed Status (SMOTE)",
  cex           = 0.8,
  split.cex     = 0.9,
  nn.cex        = 0.8,
  tweak         = 0.9,
  faclen        = 0,
  digits        = 2,
  ycompress     = FALSE
)
```

### 6.3 Variable Importance

```{r var-importance, fig.width=7, fig.height=4}
importance_df <- as.data.frame(tree_model$variable.importance) |>
  rownames_to_column("Variable") |>
  rename(Importance = `tree_model$variable.importance`) |>
  arrange(desc(Importance))

print(importance_df)

ggplot(importance_df, aes(x = reorder(Variable, Importance),
                           y = Importance, fill = Variable)) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_fill_brewer(palette = "Set2") +
  labs(
    title = "Variable Importance - Decision Tree",
    x = "Feature", y = "Importance"
  ) +
  theme_minimal()
```

### 6.4 Evaluate on Original Data

```{r tree-eval}
armedpredict$tree_pred <- factor(
  predict(tree_model,
          newdata = armedpredict[, feat_cols],
          type    = "class"),
  levels = c("No", "Yes")
)

tree_prob <- predict(tree_model,
                     newdata = armedpredict[, feat_cols],
                     type    = "prob")[, "Yes"]

tree_conf <- confusionMatrix(
  data      = armedpredict$tree_pred,
  reference = armedpredict$armed,
  positive  = "Yes"
)
print(tree_conf)

roc_tree <- roc(armedpredict$armed_yn, tree_prob)
auc_tree  <- auc(roc_tree)
cat("\nDecision Tree AUC:", round(auc_tree, 4), "\n")
```

---

## 7. Model Comparison

```{r model-comparison}
comparison <- data.frame(
  Model         = c("Logistic Regression", "Decision Tree"),
  Training_Data = rep("SMOTE-balanced", 2),
  Accuracy      = c(
    round(conf_mat$overall["Accuracy"],      4),
    round(tree_conf$overall["Accuracy"],     4)
  ),
  Sensitivity   = c(
    round(conf_mat$byClass["Sensitivity"],   4),
    round(tree_conf$byClass["Sensitivity"],  4)
  ),
  Specificity   = c(
    round(conf_mat$byClass["Specificity"],   4),
    round(tree_conf$byClass["Specificity"],  4)
  ),
  AUC           = c(
    round(auc_val,  4),
    round(auc_tree, 4)
  )
)

knitr::kable(comparison,
             caption = "Model Performance Comparison (evaluated on full original data)")
```

```{r roc-overlay, fig.width=8, fig.height=5}
ggroc(
  list(
    "Logistic Regression" = roc_obj,
    "Decision Tree"       = roc_tree
  ),
  size = 1
) +
  geom_abline(slope = 1, intercept = 1, linetype = "dashed", color = "grey50") +
  scale_color_manual(values = c("steelblue", "darkorange")) +
  labs(
    title = "ROC Curve Comparison - Logistic Regression vs Decision Tree",
    x     = "Specificity",
    y     = "Sensitivity",
    color = "Model"
  ) +
  theme_minimal()
```

---

## 8. Interpretation

**Logistic Regression:** None of the four predictors reached statistical
significance (all p-values > 0.23), and the AUC of 0.539 is only marginally
above the random baseline of 0.50. The predicted probability plots show
directionally sensible but practically negligible trends — higher poverty
and unemployment associate with slightly higher predicted probability of
being armed, while higher income and college attainment associate with
slightly lower probability. However, all predictions fall within a narrow
band (~0.45–0.65), insufficient for meaningful classification.

**Decision Tree:** With an AUC of 0.740, the tree appears to outperform
logistic regression substantially. However, this result should be interpreted
with caution: because the model is evaluated on the same data it was
effectively trained on (the original data is a subset of the SMOTE-augmented
training set), this figure is likely inflated by overfitting. The deep tree
structure (27 splits) further suggests the model has memorized patterns in
the balanced training data that do not generalize.

**Overall:** The consistent finding across both methods is that tract-level
socioeconomic features carry limited predictive signal for whether a victim
was armed. Whether a person carries a weapon appears to be driven more by
individual-level factors — such as age, personal history, or situational
context — than by the economic characteristics of the neighborhood where
the incident occurred.
